---
title: "Data 607 - Assignment 3 - Working with Stringr"
author: "Habib Khan"
date: "September 14, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to problems 3 and 4 from chapter 8 of Automated Data Collection in R.  Problem 9 is extra credit.  You may work in a small group, but please submit separately with names of all group participants in your submission.
Here is the referenced code for the introductory example in #3:
 raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
 
 
 
```{r}
# Open the required libraries
library(stringr)
library(png)
```
 
For the sake of clarity let's import the png file here to make it visible in png form. 

```{r}
img <- readPNG('week3_problems.png')
grid::grid.raster(img)
```

Copy the introductory example and store it in extracted name and extract vector name

```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
names
```



# Problem 3.1

Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

```{r}
# Sort the names

name1 <- str_replace_all(names, "(.+)(, .+)$", "\\2 \\1")
name1
```

```{r}
# Remove commas if any

name2 <- str_replace_all(name1, ", ", "")
name2
```

```{r}
# Now let's remove all prefex from names

name3 <- str_replace_all(name2, "[A-Z][a-z]([a-z]?)\\.", " ")
name3
```

# Problem 3.2

This is list of previous unedited names which has titles and prefex. 
```{r}
df1 <- data.frame(names)
df1$title <- str_detect(string= names, pattern= "\\w{2,3}\\.")
df1
```

Now let's check the final vector name3 in which all prefex and unnecessary things had removed.

```{r}
df2 <- data.frame(name3)
df2$title <- str_detect(string= name3, pattern= "\\w{2,3}\\.")
df2
```

All the prefex and titles has been removed as shown above. 

# Problem 3.3

```{r}
df2$secondname <- str_detect(string = name3, pattern= "[A-Z]{1}\\.")
df2
```



# Problem 4

```{r}
# 4.1 - This code helps bring out the numbers along with $

vector1 <- c("dsd28943889ef$asdffd", "sdf3323222$899", "sdfdsf3893$sdf")
test41 <- unlist(str_extract_all(vector1, pattern= "[0-9]+\\$" ))
test41
```

```{r}
# 4.2 - This code helps identifying characters that have 4 characters

v2 <- c("Hello What's up", "are you coming to take class today", "We have 2 classes!", "322 842 9000")
str_extract_all(v2, pattern="\\b[a-z]{1,4}\\b")
```

```{r}
# 4.3 - This code helps extracting .txt

v3 <- c("abc.txt", "dfe.txt", "444", "ewe.txt", "422.txt")
str_extract_all(v3, pattern=".*?\\.txt$")
```

```{r}
# Problem 4.4 - This code helps identifying dates as shown below

v4 <- c("09/14/2019", "07/22/2018", "as/sd/3323", "as/23/322s")
str_extract_all(v4, pattern="\\d{2}/\\d{2}/\\d{4}")
```

```{r}
# Problem 4.5 - This code works to extract html codes as shown below 

v5 <-c("<asdfasdf 322332 sdd>", "<aasfd HTML 233223>", "<HTML FILE> CHECK ME OUT </HTML FILE>", "<HTML FILE>")
str_extract_all(v5, pattern= "<(.+?)>.+?</\\1>")
```
```{r}
# Problem 9

encryptedcode ="clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
regex="[[:upper:].]+"

str_replace_all(paste(unlist(str_extract_all(encryptedcode, regex)), collapse=""), pattern="[\\.]+", replacement= " ")
```

