---
title: "Data 607 - Assignment 2 / Importing Data from MySQL"
author: "Habib Khan"
date: "September 8, 2019"
output:
  html_document: default
  pdf_document: default
---


```

Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results (observations) and store them in a SQL database. Load the information into an R dataframe.
Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.


# Load the necessary libraries

```{r echo=FALSE}

library(dbplyr)
library(RMySQL)
library(dplyr)
library(knitr)
```

# Connect RStudio with MySQL database to load the file

```{r}
mydb = dbConnect(MySQL(),
                 user='root', password='password',
                 dbname='607_hw2', host='localhost')
```


# Retrieve the joined tables together through Join Query

```{r}
mysql <- "select movies.movieid, movies.movieyear, respondent1.ratingid, respondent1.John, respondent1.James, 
respondent1.Peter, respondent1.Carol, respondent1.Steve from movies 

left join respondent1

on movies.movieid = respondent1.movieid;"

moviename <- dbGetQuery(mydb, mysql)
moviename
```

