---
title: "Data 607 - Project 1"
author: "Habib Khan and Farhana Zahir"
date: "September 21, 2019"
output: 
  html_document:
    toc:   true
    toc_float:  true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

If you have questions about the meaning of the data or the results, please post them on the discussion forum. Data science, like chess, is a game of back and forth…

The chess rating system (invented by a Minnesota statistician named Arpad Elo) has been used in many other contexts, including assessing relative strength of employment candidates by human resource departments.

All of your code should be in an R markdown file (and published to rpubs.com); with your data accessible for the person running the script.

# Objective

The objective of this project is import the text file and present it in a sophisticated way as the file is given in text form and is very unstructured

```{r}
# Loading the necessary libraries
library(stringr)
library(knitr)
```


```{r}
# Reading the file
chess <- read.delim("chessinfo.txt", header= FALSE, stringsAsFactors=FALSE)
head(chess)
```

# Data cleaning Process

```{r}
chess2  <- chess[-c(1:4),]
chess2 %>% head()
```

```{r}
#  Check the file if it is character
class(chess2)
```

```{r}
# There are lines after every two rows which needs to be eliminated. Let's do that 

name <- chess2[seq(from=1, to = length(chess2), by=3)]
ratings <- chess2[seq(from=2, to = length(chess2), by=3)]
```

```{r}
# Extracting relevant data
new_id <- as.integer(str_extract(name, "\\d+"))
new_name <- str_trim(str_extract(name, "(\\w+\\s){2,3}"))
new_point <- as.numeric(str_extract(name, "\\d.\\d"))
new_rate <- as.integer(str_extract(str_extract(ratings, "\\D\\d{3,4}\\D"), "\\d+"))
new_state <- str_extract(ratings, "\\w\\w")
new_oppid <- str_extract_all(str_extract_all(name, "\\d+\\|"), "\\d+")
```

```{r}
# Now let's convert the lists into df
op1 <- sapply(new_oppid, length)
max <- seq_len(max(op1))
op2 <- t(sapply(new_oppid, "[", i=max))
df2 <- as.data.frame(op2)


# Now let's convert these with ratings

v <- data.frame(new_id, new_rate)
df3 <- df2
df3[] <- v$new_rate[match(unlist(df3), v$new_id)]
head(df3)
```

```{r}
# Now let's calculate the average to print the final table afterwards

df3$opponentavg <- round(apply(df3,1,mean, na.rm=TRUE))
head(df3)
```

```{r}
# We are done with data cleaning, let's print the final table 
new_table <- data.frame(new_id, new_name, new_state, new_point, new_rate, df3$opponentavg)

# Table above is finalized but let's rename the column's name to make it more sophisticated

colnames(new_table) <- c("ID", "Player_Name", "State", "Point", "PreRating", "OpponentRating")
kable(new_table)
```


```{r}
# Convert the file into csv

write.table(new_table, file = "chessratings.csv", row.names=FALSE,sep = ",")
```

# Analysis and Visualization

*Player's Pre Ratings compared to Opponent's Average Ratings*

There seems to be a slight positive association between each player's ratings and the opponent's average pre rating. However, the R square is too low for any significant association.

```{r}
reg1<-lm(PreRating~ OpponentRating, data=new_table)
summary(reg1)
plot(new_table$`PreRating`, new_table$`OpponentRating`, xlim=c(800,1800), ylim=c(1000,1800), main="PreRating vs Oppononent avg Rating", xlab="Player PreRating", ylab="Opponent Average PreRating", abline(reg1))
```

*Pre Ratings statistics by State*

We can see that highest number of players in the dataset are from MI and the lowest number from OH. Players from OH have a much higher mean Pre-ratings score compared to the other two states.

```{r}
boxplot(PreRating~State, data = new_table, xlab = "State",
   ylab = "Preratings by State", main = "Statewise Average Player Pre-ratings")
```

